Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2020-08-12 docs/00-manuscripts/ieee/ieee-manuscript.tex
remove jannes inline comments
fix table width of tuning ranges table using tabularx
update keywords

\usepackage{booktabs}
\usepackage{multirow}
\usepackage{adjustbox}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
	This study analyzed highly-correlated, feature-rich datasets from hyperspectral remote sensing data using multiple machine and statistical-learning methods.
	The effect of filter-based feature-selection methods on predictive performance was compared.
	Also, the effect of multiple expert-based and data-driven feature sets, derived from the reflectance data, was investigated.
	Defoliation of trees (\%) was modeled as a function of reflectance, and variable importance was assessed using permutation-based feature importance.
% JM: I wouldn't indicate abbreviations in the abstract if you don't use them anymore within the abstract
	Overall support vector machine (SVM) outperformed others such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
	The combination of certain feature sets showed small increases in predictive performance while no substantial differences between individual feature sets were observed.
	For some combinations of learners and feature sets, filter methods achieved better predictive performances than the unfiltered feature sets, while ensemble filters did not have a substantial impact on performance.

@@

% Note that keywords are not normally used for peer-review papers.
\begin{IEEEkeywords}
	hyperspectral imagery, forest health monitoring, machine learning, feature selection, feature effects, model comparison, filter, imaging spectroscopy
\end{IEEEkeywords}

@@

% Ensemble filter methods
Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' can enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017a}.
This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses the idea of stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
% JM: I have no idea what the Borda ensemble filter is. Maybe add one more sentence on this specific filter
In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
Its final feature order is determined by the sum of all single filters ranks.

% Ensuring a fair weighting in the ensemble
% JM: reading further on, it might be that there are only three classes, please refine, Patrick!
Filter methods can be grouped into groups which are formed out of three binary classes: multivariate or univariate feature use, correlation or entropy-based importance weighting and linear and non-linear filter methodology.
Care needs to be taken to not weigh certain classes more than others in the ensemble as otherwise the final ranking result will be biased.
In this study this was taken care of by checking the rank correlations (Spearman's correlation) of the generated feature rankings of all methods against each other.
If filter pairs showed a correlation of 0.9 or higher, only one of the two was included into the ensemble filter, selected at random.
This ensured that the ensemble filter composition was not biased towards a certain group of filter methods.

@@

The idea of splitting the features into different sets originated from the question whether feature-engineered indices derived from reflectance values have a positive effect on model performance.
\cite{pena2017} is an exemplary study which used this approach in a spectro-temporal setting.
% JM: just a little reminder to avoid sentences where the main verb appears long after the sentence has started
Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.
However, rather than only looking at these three groups also combinations of such were taken into account:

@@

% parameter limits
\begin{table}[b!]
\centering
\caption[t]{Hyperparameter ranges and types for each model.
	Hyperparameter notations from the respective R packages were used.}
\begingroup\footnotesize
\begin{tabular}{llllrrr}
	\\
	\specialcell{Model                                          \\ (package)}                & Hyperparameter              & Type    & Start     & End      & Default \\
	\toprule
	\multirow{3}{*}{\specialcell{RF                                 \\ (ranger)}}       & \texttt{$x_try$}    & dbl & 0         & 0.5      & -       \\
	 & \texttt{min.node.size}      & int & 1        & 10      & 1   \\
	 & \texttt{sample.fraction}    & dbl & 0.2      & 0.9     & 1   \\
	\midrule
	\multirow{2}{*}{\specialcell{SVM                                \\ (kernlab)}}     & \texttt{C}                  & dbl & $2^{-10}$ & $2^{10}$ & 1       \\
	 & \texttt{$\sigma$}           & dbl & $2^{-5}$ & $2^{5}$ & 1   \\
	\midrule
	\multirow{7}{*}{\specialcell{XGBoost                            \\ (xgboost)}} & \texttt{nrounds}            & int & 10        & 600      & -       \\
	 & \texttt{colsample\_bytree}  & dbl & 0.3      & 0.7     & 1   \\
	 & \texttt{subsample}          & dbl & 0.25     & 1       & 1   \\
	 & \texttt{max\_depth}         & int & 1        & 10      & 6   \\
	 & \texttt{gamma}              & int & 0        & 10      & 0   \\
	 & \texttt{eta}                & dbl & 0.01     & 0.6     & 0.3 \\
	 & \texttt{min\_child\_weight} & int & 0        & 20      & 1   \\
	\bottomrule
\end{tabular}
\endgroup
\label{tab:hyperparameter_limits}
\end{table}

>>>>

\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{adjustbox}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
	This study analyzed highly-correlated, feature-rich datasets from hyperspectral remote sensing data using multiple machine and statistical-learning methods.
	The effect of filter-based feature-selection methods on predictive performance was compared.
	Also, the effect of multiple expert-based and data-driven feature sets, derived from the reflectance data, was investigated.
	Defoliation of trees (\%) was modeled as a function of reflectance, and variable importance was assessed using permutation-based feature importance.
	Overall support vector machine (SVM) outperformed others such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
	The combination of certain feature sets showed small increases in predictive performance while no substantial differences between individual feature sets were observed.
	For some combinations of learners and feature sets, filter methods achieved better predictive performances than the unfiltered feature sets, while ensemble filters did not have a substantial impact on performance.

@@

% Note that keywords are not normally used for peer-review papers.
\begin{IEEEkeywords}
	hyperspectral imagery, forest health monitoring, machine learning, feature selection, feature effects, model comparison
\end{IEEEkeywords}

@@

% Ensemble filter methods
Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' can enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017a}.
This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses the idea of stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
Its final feature order is determined by the sum of all single filters ranks.

% Ensuring a fair weighting in the ensemble
Filter methods can be grouped into groups which are formed out of three binary classes: multivariate or univariate feature use, correlation or entropy-based importance weighting and linear and non-linear filter methodology.
Care needs to be taken to not weigh certain classes more than others in the ensemble as otherwise the final ranking result will be biased.
In this study this was taken care of by checking the rank correlations (Spearman's correlation) of the generated feature rankings of all methods against each other.
If filter pairs showed a correlation of 0.9 or higher, only one of the two was included into the ensemble filter, selected at random.
This ensured that the ensemble filter composition was not biased towards a certain group of filter methods.

@@

The idea of splitting the features into different sets originated from the question whether feature-engineered indices derived from reflectance values have a positive effect on model performance.
\cite{pena2017} is an exemplary study which used this approach in a spectro-temporal setting.
Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.
However, rather than only looking at these three groups also combinations of such were taken into account:

@@

% parameter limits
\begin{table}[b!]
\centering
\caption[t]{Hyperparameter ranges and types for each model.
	Hyperparameter notations from the respective R packages were used.}
\begingroup\footnotesize
\begin{tabularx}{.5\textwidth}{llXXrXX}
	\\
	\specialcell{Model                                          \\ (package)}                & Hyperparameter              & Type    & Start     & End      & Default \\
	\toprule
	\multirow{3}{*}{\specialcell{RF                                 \\ (ranger)}}       & \texttt{$x_try$}    & dbl & 0         & 0.5      & -       \\
	 & \texttt{min.node.size}      & int & 1        & 10      & 1   \\
	 & \texttt{sample.fraction}    & dbl & 0.2      & 0.9     & 1   \\
	\midrule
	\multirow{2}{*}{\specialcell{SVM                                \\ (kernlab)}}     & \texttt{C}                  & dbl & $2^{-10}$ & $2^{10}$ & 1       \\
	 & \texttt{$\sigma$}           & dbl & $2^{-5}$ & $2^{5}$ & 1   \\
	\midrule
	\multirow{7}{*}{\specialcell{XGBoost                            \\ (xgboost)}} & \texttt{nrounds}            & int & 10        & 600      & -       \\
	 & \texttt{colsample\_bytree}  & dbl & 0.3      & 0.7     & 1   \\
	 & \texttt{subsample}          & dbl & 0.25     & 1       & 1   \\
	 & \texttt{max\_depth}         & int & 1        & 10      & 6   \\
	 & \texttt{gamma}              & int & 0        & 10      & 0   \\
	 & \texttt{eta}                & dbl & 0.01     & 0.6     & 0.3 \\
	 & \texttt{min\_child\_weight} & int & 0        & 20      & 1   \\
	\bottomrule
\end{tabularx}
\endgroup
\label{tab:hyperparameter_limits}
\end{table}
